Optimizing Large Scale Surveys for Photometric Calibration, \\ The Euclid Dark Energy Mission as an Example

### 2011-07-03 Documentation/cross_cal_paper.tex
euclid: paper update, plotting function update: nicer figures
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@18943 03a3cea6-2c03-0410-ac9b-9271b1c66c29

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{coverage.png}
\end{center}
\caption{An example catalog of fake sources used within the simulations. The sources are shown at their position on a portion of the sky (left) and as a histogram of their magnitudes (right).
\label{fig:sky}}
\end{figure}

@@

\begin{displaymath}
\sigma_{{act}}^{2} = (1 + \epsilon) \,\alpha^{2} + \eta^{2}\, c^{2} 
\end{displaymath}
\noindent{}where $\epsilon$ is a random number, in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$. An example of the assumed compared to the true uncertainty variances is shown in Fig. \ref{fig:count_uncertainty}.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{invvar_plot.png}
\end{center}
\caption{The assumed and the actual variances as a function of counts.\label{fig:count_uncertainty}}
\end{figure}

@@

\subsubsection{Instrument Response Refinement: \textbf{\textit{p\_step}}}
The instrument response parameters can now be refined with the latest star flux estimates. To do this we minimize the total error function for all measurements with respect to the instrument response parameters.

@@

\section{Results}
\subsection{Simple Survey Strategies}
In this section we consider four simple, very different, survey strategies.

\subsection{Surveys Strategies}
We consider two extreme survey examples. The first, called Strategy A, consists of a regular grid of pointings over the sky with a small overlap between neighboring fields. Nine passes are performed over the sky and each time the grids are aligned exactly. The second survey, Strategy D, has the same number of fields as Strategy A, but the pointing and orientation of these fields are random. These two surveys can be seen in Fig. \ref{fig:surveys}.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{coverage.png}
\end{center}
\caption{Left: The A Strategy has regular pointings over the sky with a small overlap between adjacent fields. Nine passes are performed over the sky and each time the grids are aligned exactly. Right: The D Strategy has the same number of pointings as the A Strategy, but the pointings and orientations are random.\label{fig:surveys}}
\end{figure}

The coverage of each strategy is shown in Fig. \ref{fig:coverage}.

\subsection{Ubercalibration Performance}

The difference in ubercalibration method's ability to fit the flat field and the star magnitudes for the two different Strategies is striking. With the random Strategy (D) we can get very close to the actual flat field in a small number of ubercalibration iterations (see Fig \ref{fig:D}). The well ordered Strategy (A), on the other hand, does not converge to the correct flat field even with 1,000 ubercal iterations. Even when we start with almost the correct flat-field parameters (those found with the D Strategy - shown in Fig. \ref{fig:D}), the ubercalibration procedure does not converge. In fact, with each iteration solution drifts away from the actual flat-field. Fig \ref{fig:A} shows the fitted flat-field from Strategy A after 1,000 with the initial parameters shown in Fig \ref{fig:D}.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{D_15_ff.png}
\end{center}
\caption{The fitted flat field compared to the actual flat field found with Strategy D. \label{fig:D}}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{A_1000_ff.png}
\end{center}
\caption{The fitted flat field compared to the actual flat field calculated found with Strategy A. The ubercalibration procedure was started with almost exact flat field parameters (those found from Strategy D - see Fig. \ref{fig:D}) and has drifted away from the actual solution. \label{fig:A}}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.47\textwidth]{coverage.png}
\end{center}
\caption{The coverage of both of the surveys investigated. \label{fig:coverage}}
\end{figure}

\section{Discussion}
small changes to survey that greatly improve coverage plots
random strategy can be defined so that the telescope does not make large steps in pointing, orientation
have't addressed time dependence, how do these surveys with time will be different for different assumptions on temporal variability. 
is the small scale power we put in realistic?
- survey D is much better than survey C for astrometric self calibration 
we haven't thought about gaps
\end{document}

>>>>

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{survey.png}
\end{center}
\caption{An example catalog of fake sources used within the simulations. The sources are shown at their position on a portion of the sky (left) and as a histogram of their magnitudes (right).
\label{fig:sky}}
\end{figure}

@@

\begin{displaymath}
\sigma_{{act}}^{2} = (1 + \epsilon) \,\alpha^{2} + \eta^{2}\, c^{2} 
\end{displaymath}
\noindent{}where $\epsilon$ is a random number, in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$. An example of the assumed compared to the true uncertainty variances is shown in Fig. \ref{fig:invvar}.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{invvar.png}
\end{center}
\caption{This plot highlights the difference between the assumed (solid line) and the actual (black points) measurement uncertainty variances.\label{fig:invvar}.}
\end{figure}

@@

\subsubsection{Instrument Response Refinement (\textit{p\_step})}
The instrument response parameters can now be refined with the latest star flux estimates. To do this we minimize the total error function for all measurements with respect to the instrument response parameters.

@@

\section{Results}
\subsection{Simple Survey Strategies}
In this section we consider four simple, very different, survey strategies. We investigate how well the cross-calibration procedure can constrain the relative instrument response with these surveys. We label these strategies A to D. The different strategies are summarized in Tab. \ref{tab:surveys} and are shown in Fig. \ref{fig:surveys}. Strategy A is the simplest strategy; the field is regularly tiled with a small (XX\%) overlap between adjacent pointings. The pointings in the 12 passes over the same field are exactly aligned. Survey B is the same as A, but with each pass over the field the orientation of the telescope is rotated by $30^\circ$. Survey C is more complex. The first pass over the field is the same as in Survey A, with $12 \times 12$ pointings. In the next pass, one of the pointings in the $\alpha$ directions is removed and one is added in the $\beta$, so our grid of pointings is now $13 \times 11$. The overlap in now XX\% and XX\% in the $\alpha$ and $\beta$ directions respectively. In the third pass over the field, the change to the point grid, compared to the first pass, is reversed and we image the field on a $11 \times 13$ grid. These three passes are then repeated four times. Survey D is quasi-random. The orientation of each pointing is random. The pointing positions are quasi-random: the pointing center is randomly assigned within a $0.7 \times 0.7$ box whose center lies on the pointing grid in Survey A. This is done to ensure a uniform coverage of the field.

\begin{table}
\begin{tabular}{|c|p{5.5cm}|p{3cm}|p{3cm}|}
\hline \textbf{Survey Name} & \textbf{Pointing Center} & \textbf{Pointing Orientation} & \textbf{Number of Pointings}\\ 
\hline \textbf{A} & Uniform Grid ($12\times12$) & Constant & 1728\\ 
\hline \textbf{B} & Uniform Grid ($12\times12$) & 30$^\circ$ Rotation Between Passes & 1728 \\ 
\hline \textbf{C} & Pass 1:  Uniform Grid ($12\times12$) Pass 2:  Uniform Grid ($1§\times11$) Pass 3:  Uniform Grid ($13\times11$)& Constant & 1720 \\ 
\hline \textbf{D} & Quasi-random & Random & 1728 \\ 
\hline 
\end{tabular}   
\caption{A summary of the four simple survey strategies investigated.\label{tab:surveys}}
\end{table}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{simple_surveys.png}
\end{center}
\caption{The four simple survey strategies described in the text and summarized in Tab. \ref{tab:surveys}. Surveys A, B and D have 1728 pointings and survey C has 1720 pointings.\label{fig:surveys}}
\end{figure}

\begin{table}
\begin{tabular}{|c|p{4cm}|c|c|}
\hline Survey & Number of Iterations Until Converges & Instrument Response Fit Badness & RMS Source Error \\ 
\hline A & no convergence & - & - \\ 
\hline B & no converges & - & - \\ 
\hline C &  &  &  \\ 
\hline D &  &  &  \\ 
\hline 
\end{tabular}
\caption{The performance of the cross-calibration of the datasets with the survey strategies detailed in Tab. \ref{tab:surveys}\label{fig:simple_results}}
\end{table}

We can draw a strong conclusion from these simple survey strategies: small and regular overlaps between adjacent pointings is essentially useless for this kind of cross-calibration. The simple strategies, where the sky is just tiled, should be avoided in large scale imaging surveys.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{fit_example.png}
\end{center}
\caption{The fitted flat field compared to the actual flat field found with Strategy D. \label{fig:D}}
\end{figure}

\end{document}
